---
title: "The Analytical Policeman"
output: html_notebook
---

### Loading data

```{r}
mvt = read.csv("mvt.csv", stringsAsFactors = FALSE)
str(mvt)
```
### Coverting to date-time object

```{r}
mvt$Date = strptime(mvt$Date, format = "%m/%d/%y %H:%M")
```

```{r}
mvt$Weekday = weekdays(mvt$Date)
mvt$Hour = mvt$Date$hour
```

```{r}
str(mvt)
```

### Creating lineplots

```{r}
weekday_counts = as.data.frame(table(mvt$Weekday))
str(weekday_counts)
```

```{r}
library(ggplot2)
```

```{r}
weekday_counts$Var1 = factor(weekday_counts$Var1, ordered = TRUE, levels =  c("Sunday","Monday","Tuesday","Wednesday","Thursday","Friday","Saturday"))
```

```{r}
ggplot(weekday_counts, aes(x = Var1, y = Freq)) + geom_line(aes(group = 1)) + 
  xlab("Day of the week") + ylab("Total motor vehicle thefts")
```

```{r}
table(mvt$Weekday, mvt$Hour)
DayHourCounts = as.data.frame(table(mvt$Weekday, mvt$Hour))
```

```{r}
str(DayHourCounts)
head(DayHourCounts)
DayHourCounts$Hour = as.numeric(as.character(DayHourCounts$Var2))
```

```{r}
ggplot(aes(x = Hour, y = Freq), data = DayHourCounts) + 
  geom_line(aes(group = Var1, color = Var1), size = 2)
```

### Creating heatmaps

* Reordering var1

```{r}
DayHourCounts$Var1 = factor(DayHourCounts$Var1, ordered = TRUE, levels = c("Monday","Tuesday","Wednesday","Friday","Saturday","Sunday"))
```

```{r}
ggplot(aes(x = Hour, y = Var1), data = DayHourCounts) + geom_tile(aes(fill = Freq)) +
  scale_fill_gradient(name = "Total MV thefts", low = "white", high = "red") + 
  theme(axis.title.y = element_blank())
```

### A geographical hot spot map

```{r}
library(maps)
library(ggmap)
```
* Loading map of Chicago

```{r}
chicago = get_map(location = "chicago", zoom = 11)
```

```{r}
ggmap(chicago)
```

* Plotting first 100 motor thefts on the map

```{r}
ggmap(chicago) + geom_point(data = mvt[1:100,], aes(x = Longitude, y = Latitude))
```

* Rounding latitudes and longitudes to two digits and creating crime counts data

```{r}
LatLonCounts = as.data.frame(table(round(mvt$Latitude,2), round(mvt$Longitude,2)))
str(LatLonCounts)
summary(LatLonCounts$Freq)
```

```{r}
LatLonCounts$Lat = as.numeric(as.character(LatLonCounts$Var1))
LatLonCounts$Long = as.numeric(as.character(LatLonCounts$Var2))
```

```{r}
summary(LatLonCounts)
str(LatLonCounts)
```

* Plotting above points on map making the size and color of points depend upon the total number  
of motor vehicle thefts

```{r}
ggmap(chicago) +
  geom_point(aes(x = Long, y = Lat, color = Freq, size = Freq), data = LatLonCounts) +
  scale_color_gradient(low = "yellow", high = "red")
```

* Plotting heatmaps

```{r}
ggmap(chicago) + 
  geom_tile(data = LatLonCounts, aes(x = Long, y = Lat, alpha = Freq), fill = "red")
```

Note: We see that in the above maps points are squares are plotted in the water, this corresponds  
to the area in water, we can remove that by choosing points having frequency > 0

* Choosing points with frequency greater than 0

```{r}
LatLonCounts2 = subset(LatLonCounts, LatLonCounts$Freq > 0)
str(LatLonCounts2)
```

```{r}
ggmap(chicago) + 
  geom_tile(aes(x = Long, y = Lat, alpha = Freq), data = LatLonCounts2, fill = "red")
```

