---
title: "Basic scatterplot using ggplot"
output: html_notebook
---

### Loading dataset

```{r}
WHO = read.csv("WHO.csv")
str(WHO)
```

### Scatter plot => GNI vs Fertility rate

```{r}
plot(WHO$GNI, WHO$FertilityRate)
```

* Using ggplot

```{r}
library(ggplot2)
```

```{r}
scatterplot = ggplot(WHO, aes(x = GNI, y = FertilityRate))
scatterplot + geom_point()
```

* linegraph

```{r}
scatterplot + geom_line()
```

* scatterplot with blue triangles

```{r}
scatterplot + geom_point(color = "blue", size = 3, shape = 17)
```

* scatterplot with stars

```{r}
scatterplot + geom_point(color = "darkred", size = 3, shape = 8)
```

* Adding title to plot

```{r}
scatterplot + geom_point(color = "darkred", size = 3, shape = 15) + 
  ggtitle("Fertility rate vs Gross national income")
```

* saving plot to file

```{r}
fertilityGNIplot = scatterplot + geom_point(color = "darkred", size = 3, shape = 8) + 
  ggtitle("Fertility rate vs Gross national income")
```

```{r}
pdf("Myplot.pdf")
print(fertilityGNIplot)
dev.off()
```

### Advance scatterplots using ggplot

* Coloring the points by region

```{r}
ggplot(WHO, aes(x = GNI, y = FertilityRate, color = Region)) +
  geom_point()
```

* Coloring the points by life expectancy

```{r}
ggplot(WHO, aes(x = GNI, y = FertilityRate, color = LifeExpectancy)) +
  geom_point()
```

* Fertility rate vs Under15

Note: In the below graph we see that the as the fertility rate increases, Under15 starts increasing  
less.
   
```{r}
ggplot(WHO, aes(x = FertilityRate, y = Under15)) + geom_point()
```
Note: We see that there is a linear relationship between the log(Fertilityrate) and Under15 variable  

```{r}
ggplot(WHO, aes(x = log(FertilityRate), y = Under15)) + geom_point()
```

* Building linear regression model between log(FertilityRate) and Under15

```{r}
model1 = lm(Under15 ~ log(FertilityRate), data = WHO)
summary(model1)
```

* Adding regression line to the plot

Note: The shaded portion around the regression line is the 99% confidence interval

```{r}
ggplot(WHO, aes(x = log(FertilityRate), y = Under15)) +
  geom_point() + stat_smooth(method = "lm", level = 0.99, color = "orange")
```

* FertitlityRate vs Under15 by region

```{r}
ggplot(WHO, aes(x = FertilityRate, y = Under15, color = Region)) + 
  geom_point()
```

