---
title: "Heat map on United States"
output: html_notebook
---

### Loading required libraries

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggmap)
```

### Loading dataset

```{r}
murders = read.csv("murders.csv")
str(murders)
```

### Loading map of United States

```{r}
statesMap = map_data("state")
str(statesMap)
```

```{r}
ggplot(data = statesMap, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black")
```

### Making sure that the state name in our murders dataset is same as in statesMap

```{r}
murders$region = tolower(murders$State)
```

### Joining the dataframes "statesMap" and "murders"

```{r}
murderMap <- merge(statesMap, murders, by = "region")
str(murderMap)
```

### Plotting the murders on the map

* We can see that states of California and Texas have unusually high number of murders, but that may be because of high population, so we'll plot a map for population.

```{r}
ggplot(aes(x = long, y = lat, group = group, fill = Murders), data = murderMap) +
  geom_polygon(color = "black") + 
  scale_fill_gradient(low = "black", high = "red", guide = "legend")
```

* Plotting a map of population of each state

```{r}
ggplot(aes(x = long, y = lat, group = group, fill = Population), data = murderMap) +
  geom_polygon(color = "black") + 
  scale_fill_gradient(low = "black", high = "red", guide = "legend")
```

* Plotting a map for murder rate variable

Note: Our map for murder rate is maroon in color for most parts, there is not any red state, because Washington,DC is an outlier.

```{r}
murderMap$MurderRate <- murderMap$Murders / murderMap$Population * 100000
```

```{r}
ggplot(aes(x = long, y = lat, group = group, fill = MurderRate), data = murderMap) +
  geom_polygon(color = "black") + 
  scale_fill_gradient(low = "black", high = "red", guide = "legend")
```

* Removing Washington DC from data

```{r}
ggplot(aes(x = long, y = lat, group = group, fill = MurderRate), data = murderMap) +
  geom_polygon(color = "black") + 
  scale_fill_gradient(low = "black", high = "red", guide = "legend", limits = c(0,10))
```

* States with highest number of people owning guns

```{r}
ggplot(aes(x = long, y = lat, group = group, fill = GunOwnership), data = murderMap) +
  geom_polygon(color = "black") + 
  scale_fill_gradient(low = "black", high = "red", guide = "legend")
```

