---
title: "MRI image"
output: html_notebook
---

### Reading the data

```{r}
healthy = read.csv("healthy.csv", header = FALSE)
```

### Creating matrix

```{r}
healthyMatrix = as.matrix(healthy)
str(healthyMatrix)
```

### Generating the image

```{r}
image(healthyMatrix, axes = FALSE, col = grey(seq(0,1,length = 256)))
```

### Creating a vector

```{r}
healthyVector = as.vector(healthyMatrix)
```

### Performing clustering

```{r}
str(healthyVector)
```

Note: We see that the length of vector *healthyVector* is 365636, the distance vector will have  
365636*(365635)/2 = approx 67 billion values, so we cannot create a distance vector so huge,  
and unfortunately cannot apply hierarchical clustering.


* Setting number of clusters to 5

```{r}
k = 5
set.seed(1)
```

```{r}
KMC = kmeans(healthyVector, centers = k, iter.max = 1000)
```

```{r}
KMC
str(KMC)
```

```{r}
healthyClusters = KMC$cluster
KMC$centers
```


### Generating the segmented image

```{r}
dim(healthyClusters) = c(nrow(healthyMatrix),ncol(healthyMatrix))
image(healthyClusters, axes = FALSE, col = rainbow(k))
```


### Using clustering to identify tumor in MRI

```{r}
tumor = read.csv("tumor.csv", header = FALSE)
tumorMatrix = as.matrix(tumor)
tumorVector = as.vector(tumorMatrix)
```

```{r}
library(flexclust)
```

```{r}
KMC.kcca = as.kcca(KMC, healthyVector)
```



