                                    Predicting the Baseball World Series Champion
                                    
Loading Dataset

```{r}
baseball = read.csv("baseball.csv")
nrow(baseball)
length(table(baseball$Year))
str(baseball)
```
Limiting to team making to playoffs

```{r}
baseball = subset(baseball, Playoffs == 1)
nrow(baseball)
table(table(baseball$Year))
```
Adding an imporatnt predictor

The more the number of teams in playoffs the harder it is to win the World Series. So we add a variable NumCompetitors which is equal to the
number of teams making it to the playoffs in the year of a particular team/year pair.

```{r}
PlayoffTable = table(baseball$Year)
PlayoffTable[c("1990", "2001")]
baseball$Numcompetitors = PlayoffTable[as.character(baseball$Year)]
table(baseball$Numcompetitors)

```

Bivariate models for predicting World Series Winner

```{r}
baseball$WorldSeries = as.numeric(baseball$RankPlayoffs == 1)
table(baseball$WorldSeries)
```

Making our Logistic regression model using all the independent variables

```{r}
baseballLog = glm(WorldSeries ~ Year + RA + RankSeason + Numcompetitors, data = baseball, family = "binomial")
cor(baseball[c("Year", "RA", "RankSeason", "Numcompetitors")])
```
```{r}
summary(glm(WorldSeries ~ Year, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ RA, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ RankSeason, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ Numcompetitors, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ Year + RA, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ Year + RankSeason, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ Year + Numcompetitors, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ RA + RankSeason, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ RA + Numcompetitors, data = baseball, family = binomial ))
summary(glm(WorldSeries ~ RankSeason + Numcompetitors, data = baseball, family = binomial ))
```
Note:  We can see that the AIC value is minimum for the bivariate model with Numcompetitors as independent variable, so we can say that luck matters a lot in in playoffs as claimed by Billy Beane, as number of competitors has nothing to do with quality of players.
