---
title: "An introduction to clustering"
output: html_notebook
---


###Loading data

```{r}
movies = read.table("movieLens.txt", header = FALSE, sep = "|", quote = "\"")
str(movies)
```

###Adding columns

```{r}
colnames(movies) = c("ID", "Title", "ReleaseDate", "VideoReleaseDate","IMDB","Unknown","Action","Adventure", "Animation", "Childrens",
                     "Comedy", "Crime","Documentry", "Drama","Fantasy","FilmNoir","Horror","Musical","Mystery","Romance","SciFi",
                     "Thriller","War","Western")
str(movies)
```


###Removing unwanted variables

```{r}
movies$ID = NULL
movies$ReleaseDate = NULL
movies$VideoReleaseDate = NULL
movies$IMDB = NULL
movies = unique(movies)
```

```{r}
str(movies)
```

###Movies classified as comedy

```{r}
table(movies$Comedy)
```

###Movies classifed as western

```{r}
table(movies$Western)
```

###Movies classified as romance and drama

```{r}
table(movies$Romance, movies$Drama)
```

###Calculating distances

```{r}
distances = dist(movies[2:20], method = "euclidean")
```

###Clustering movies

```{r}
clusterMovies = hclust(distances, method = "ward.D")
```

###Plotting dendogram

```{r}
plot(clusterMovies)
```


###Assigning labels to our datapoints

```{r}
clusterGroups = cutree(clusterMovies, k = 10)
```  

Note: Finding average number of movies with genre action in all the clusters. *Split* and *laaply* functions are used. The below   
command splits data into subsets based on different clusters.

```{r}
spl = split(movies[2:20], clusterGroups)
str(spl)
```

```{r}
lapply(spl,colMeans)
```

###Looking for specific films

```{r}
subset(movies, Title == "Men in Black (1997)")
```  
*Cluster for Men In Black*  

```{r}
clusterGroups[257]
```


###Dataset with movies from cluster 2

```{r}
cluster2 = subset(movies, clusterGroups == 2)
cluster2$Title[1:10]
```

```{r}
clusterGroups
```



###Clustering with k = 2

```{r}
clusterGroups2 = cutree(clusterMovies, k = 2)
```

```{r}
spl = split(movies[2:20], clusterGroups2)
```

```{r}
lapply(spl,colMeans)
```



