Letter Recognition
=========


###1. Loading and inspecting dataset

```{r}
letters = read.csv("letters_ABPR.csv")
str(letters)
summary(letters)
```

###2. Predicting just one letter


1. Just predicting whether a letter is B or not

```{r}
letters$isB = as.factor(letters$letter == "B")
```

2. Splitting dataset

```{r}
library(caTools)
set.seed(1000)
split = sample.split(letters$isB, SplitRatio = 0.5)
train = subset(letters, split == TRUE)
test = subset(letters, split == FALSE)
```

3. Accuracy of baseline model on test set

Our baseline model predicts "not B" for every data.

```{r}
table(test$isB)
cat("Accuracy of baseline model on our test set =>", (1175)/nrow(test))
```


4. Classification tree model

```{r}
library(rpart)
library(rpart.plot)
```

```{r}
CARTb = rpart(isB ~ .-letter, data = train, method = "class")
prp(CARTb)
Bpred = predict(CARTb, newdata = test, type = "class")
table(test$isB, Bpred)
cat("Accuracy of our CART model =>", (1118 + 340)/nrow(test))
```


5. Random forest model

```{r}
library(randomForest)
```

```{r}
set.seed(1000)
Bforest = randomForest(isB ~ .-letter, data = train)
BforestPred = predict(Bforest, newdata = test)
table(test$isB, BforestPred)
cat("Accuracy of random forest model =>", (1165+374)/nrow(test))
```




###3. Predicting letters A,B,P,R

1. Converting letter variable to a factor variable

```{r}
letters$letter = as.factor(letters$letter)
```

2. Splitting data

```{r}
set.seed(2000)
spl = sample.split(letters$letter, SplitRatio = 0.5)
train = subset(letters, spl == TRUE)
test = subset(letters, spl == FALSE)
table(test$letter)
cat("Baseline accuracy =>", (401)/nrow(test))
```

3. CART model

```{r}
CARTletter = rpart(letter ~ .-isB, data = train, method = "class")
prp(CARTletter)
```

```{r}
letterPred = predict(CARTletter, newdata = test, type = "class")
table(test$letter, letterPred)
cat("Accuracy of CART model =>", (348+318+363+340)/nrow(test))
```


4. Random forest model

```{r}
set.seed(1000)
letterForest = randomForest(letter ~ .-isB, data = train)
letterForestPred = predict(letterForest, newdata = test)
table(test$letter, letterForestPred)
cat("Accuracy of random forest model =>", (391+380+393+365)/nrow(test))
```

*Note: We see that the accuracy of CART model decreased significantly when transitioned to the problem of predicting of all 4 letters, but
in case of Random Forest model the decrement was not very significant.
