                         #################################################################
                         ##### Predicting parole violators using logistic regression #####
                         #################################################################


##### Loading and investigating data

```{r}
setwd("/home/amanlp7/R_datasets/")
parole = read.csv("parole.csv")
str(parole)
summary(parole)
```
##### Parolees that have violated the terms of parole

```{r}
table(parole$violator)
```
##### Converting State and Crime variables to factors

```{r}
parole$crime = as.factor(parole$crime)
parole$state = as.factor(parole$state)
summary(parole)
```
##### Splitting data into training and testing set

```{r}
set.seed(144)
library(caTools)
split = sample.split(parole$violator, SplitRatio = 0.7)
train = subset(parole, split == TRUE)
test = subset(parole, split == FALSE)
```
##### Buliding a Logistic Regression Model

```{r}
paroleLog = glm(violator ~ ., data = train, family = "binomial")
summary(paroleLog)
```
##### Evaluating the model on the test set

```{r}
predictTest = predict(paroleLog, newdata = test, type = "response")
max(predictTest)
```
##### Calculating model's senstivity and specifcity

```{r}
table(test$violator, predictTest >= 0.5)
sprintf("Sensitivity => %f", 12/23)
sprintf("Specificty => %f", 167/179)
sprintf("Accuracy => %f", 179/202)
```
##### Accuracy of base model

```{r}
table(test$violator)
sprintf("Accuracy => %f", 179/202 )
```
##### AUC value of our model

```{r}
library(ROCR)
pred = prediction(predictTest, test$violator)
as.numeric(performance(pred, "auc")@y.values)
```

