---
title: "State Data"
output: html_notebook
---

###1. Linear regression model

1. Loading dataset 

```{r}
data(state)
stateData = data.frame(state.x77)
str(stateData)
```


2. Building linear regression model

```{r}
stateLin = lm(Life.Exp ~ ., data = stateData)
summary(stateLin)
```

3. Making predictions

```{r}
statepred = predict(stateLin)
SSE = sum((statepred - stateData$Life.Exp)^2)
cat("Sum of squared errors for linear model =>", SSE)
```

4. Buliding linear regression model by taking **significant variables** from **stateLin**

```{r}
stateLin2 = lm(Life.Exp ~ Population + Frost + Murder + HS.Grad, data = stateData)
summary(stateLin2)
```


5. Making predictions using model **stateLin2**

```{r}
statepred2 = predict(stateLin2)
cat("Sum of squared errors for stateLin2 =>", sum((statepred2 - stateData$Life.Exp)^2))
```




###2.  CART models

1. Importing required libraries

```{r}
library(rpart)
library(rpart.plot)
```

2. Building CART models

```{r}
stateCART = rpart(Life.Exp ~ ., data = stateData, minbucket = 5)
prp(stateCART)
```


3. Predicting life expectancy

```{r}
stateCARTpred = predict(stateCART)
cat("Sum of squared errors for our CART model =>", sum((stateCARTpred - stateData$Life.Exp)^2))
```


4. CART model using only **Area** as an independent variable

```{r}
stateCART2 = rpart(Life.Exp ~ Area, data = stateData, minbucket = 1)
prp(stateCART2)
stateCARTpred2 = predict(stateCART2)
cat("SSE of stateCART2 =>", sum((stateCARTpred2 - stateData$Life.Exp)^2))
```

###3. Cross-validation 

1. Loading libraries

```{r}
library(caret)
library(e1071)
set.seed(111)
```

```{r}
stateControl = trainControl(method = "cv", number = 10)
cartGrid = expand.grid(.cp = seq(0.01,0.50,0.01))
stateTrain = train(Life.Exp ~ ., data = stateData, method = "rpart", trControl = stateControl, tuneGrid = cartGrid)
```
```{r}
stateTrain
```



Note: We see that cp = 0.12 is the optimal choice


2. Building CART model using cp = 0.12

```{r}
stateCART3 = rpart(Life.Exp ~ ., data = stateData, cp = 0.12)
prp(stateCART3)
stateCARTpred3 = predict(stateCART3)
cat("SSE of our tree with cp = 0.12 =>", sum((stateCARTpred3 - stateData$Life.Exp)^2))
```



Note: We observe that our CART model with cp = 0.12 has greater SSE than all other CART trees , that means , it performs slightly worse than all the other CART trees on our training set, but we'll still choose it, for our testing set, because the purpose of cross - validation is to pick a tree which will perform best on our test set, so we would expect that the model we made with the best **cp** performs best on our test set.



###4. Cross-validation for **Area** as an independent variable

1. Finding optimal value of cp

```{r}
set.seed(111)
stateControl1 = trainControl(method = "cv", number = 10)
cartGrid = expand.grid(.cp = seq(0.01,0.50,0.01))
stateTrain1 = train(Life.Exp ~ Area, data = stateData, method = "rpart", trControl = stateControl, tuneGrid = cartGrid)
```

```{r}
stateTrain1
```


2. Making model with **Area** as an independent variable with cp = 0.06

```{r}
stateCARTarea = rpart(Life.Exp ~ Area, data = stateData, cp = 0.06)
prp(stateCARTarea)
stateAreaPred = predict(stateCARTarea)
cat("SSE for our CART model using area =>", sum((stateAreaPred - stateData$Life.Exp)))
```

Note: Cross-validation is intended to select the optimal parameters for a model, it may or may not decrease the sum of squred errors, it is not intended to fit the model on dataset.