---
title: "Predicting stock return with cluster-then-predict"
output: html_notebook
---

### Loading dataset

```{r}
stocks = read.csv("StocksCluster.csv")
```

```{r}
str(stocks)
```

### Proportion of stocks having positive returns in December

```{r}
table(stocks$PositiveDec)
cat("Proportion of stocks with positive return =>",(6324)/11580)
```

### Correlation between variables

```{r}
cor(stocks)
```
### Largest mean return from January to November

```{r}
which.max(colMeans(stocks[1:11]))
```

### Smallest mean return from January to November

```{r}
which.min(colMeans(stocks[1:11]))
```

### Initial logistic regression model

```{r}
library(caTools)
set.seed(144)
```

```{r}
spl = sample.split(stocks$PositiveDec, SplitRatio = 0.7)
stocksTrain = subset(stocks, spl == TRUE)
stocksTest = subset(stocks,spl == FALSE)
```

```{r}
stocksLog = glm(PositiveDec ~ ., data = stocksTrain, family = "binomial")
summary(stocksLog)
```

```{r}
trainPred = predict(stocksLog, type = "response")
```

* Accuracy of logistic regression model on training set  

```{r}
table(stocksTrain$PositiveDec, trainPred > 0.5)
cat("Accuracy of logistic regression model on training set =>", (3640+990)/nrow(stocksTrain))
```

* Accuracy of logistic regression model on testing set

```{r}
testPred = predict(stocksLog, newdata = stocksTest, type = "response")
```

```{r}
table(stocksTest$PositiveDec, testPred > 0.5)
cat("Accuracy =>", (1553+417)/nrow(stocksTest))
```

* Accuracy on testing set of baseline model that predicts most common outcome  

```{r}
table(stocksTest$PositiveDec)
cat("Accuracy =>", (1897)/nrow(stocksTest))
```

### Clustering stocks

* The first step is to remove the dependent variable. It is done because needing to know the dependent variable value to assign an observation to a cluster defeats the purpose of the methodology.

```{r}
limitedTrain = stocksTrain
limitedTest = stocksTest
```

```{r}
limitedTrain$PositiveDec = NULL
limitedTest$PositiveDec = NULL
```

* Normalizing the variables in the training set

```{r}
library(caret)
```

```{r}
preproc = preProcess(limitedTrain)
```

```{r}
normTrain = predict(preproc, limitedTrain)
normTest = predict(preproc,limitedTest)
```
* Mean of ReturnJan variable in normTrain and normTest

```{r}
mean(normTrain$ReturnJan)
mean(normTest$ReturnJan)
```

* Setting seed for clustering

```{r}
set.seed(144)
k = 3
```

```{r}
km = kmeans(normTrain, centers = k)
```

*Number of observations in each clusters

```{r}
table(km$cluster)
```

* Obtaining training and testing set cluster assignments for our observations  

```{r}
library(flexclust)
```

```{r}
km.kcca = as.kcca(km,normTrain)
```

```{r}
clusterTrain = predict(km.kcca)
```

```{r}
clusterTest = predict(km.kcca, newdata = normTest)
table(clusterTest)
```

* Building seperate dataframes for observations in seperate clusters  

```{r}
stocksTrain1 = subset(stocksTrain, km$cluster == 1)
stocksTrain2 = subset(stocksTrain, km$cluster == 2)
stocksTrain3 = subset(stocksTrain, km$cluster == 3)
```

```{r}
stocksTest1 = subset(stocksTest, km$cluster == 1)
stocksTest2 = subset(stocksTest, km$cluster == 2)
stocksTest3 = subset(stocksTest, km$cluster == 3)
```
* Trainig set dataframe with highest average value for dependent variable  

```{r}
tapply(stocksTrain$PositiveDec, km$cluster, mean)
```

### Cluster-specific Predictions

```{r}
stocksModel1 = glm(PositiveDec ~ ., data = stocksTrain1, family = "binomial")
stocksModel2 = glm(PositiveDec ~ ., data = stocksTrain2, family = "binomial")
stocksModel3 = glm(PositiveDec ~ ., data = stocksTrain3, family = "binomial")
```
* Making predictions for each test set grouped by cluster

```{r}
PredictTest1 = predict(stocksModel1, newdata = stocksTest1, type = "response")
PredictTest2 = predict(stocksModel2, newdata = stocksTest2, type = "response")
PredictTest3 = predict(stocksModel3, newdata = stocksTest3, type = "response")
```

* Accuracy of model on test set 1

```{r}
table(stocksTest1$PositiveDec, PredictTest1 > 0.5)
cat("Accuracy =>", (101+498)/(101+103+341+498))
```

* Accuracy for model on test set 2

```{r}
table(stocksTest2$PositiveDec, PredictTest2 > 0.5)
cat("Accuracy =>", (361+862)/(361+862+301+686))
```

* Accuracy for model on test set 3

```{r}
table(stocksTest3$PositiveDec, PredictTest3 > 0.5)
cat("Accuracy =>", (69+39)/(69+19+94+39))
```

* Compute overall test-set accuracy

```{r}
AllPredictions = c(PredictTest1,PredictTest2,PredictTest3)
AllOutcomes = c(stocksTest1$PositiveDec,stocksTest2$PositiveDec, stocksTest3$PositiveDec)
```

```{r}
table(AllOutcomes, AllPredictions > 0.5)
cat("Accuracy =>", (531+1399)/(531+1399+498+1046))
```




